Reduced plasticity and relative variance in physiological rates of ectotherm populations under climate change

### 2024-05-17 ms.qmd
Rewording

Phenotypic plasticity is expected to evolve in populations when environmental variability is high but predictable and the costs of plasticity are low [@Chevin2017; @Scheiner2020; @Nunney2016; @Dewitt1998; @Reed2010]. Reversible phenotypic plasticity, such as physiological acclimation, is relatively rapid and can be fine-tuned to proximate environmental conditions making it the first ‘line-of-defense’ against rapid environmental change [@Dewitt1998; @Scheiner2020]. For example, physiological rates are known to speed up as temperature increases because of the thermodynamic effects on chemical reaction rates -- so called 'acute' temperature responses. Acute temperature increases that push organisms from their thermal optimum can be mitigated by plastically adjusting reaction rates in response to chronic temperatures through acclimation [@Seebacher2015-wi; @Havird2020-mj]. Acclimation to higher temperatures does not normally result in physiological rates returning to their original state (‘complete’ compensation - *sensu* @huey1999testing), but rather rates remain similar to acute responses at high temperatures ('no acclimation') or only partially compensate [@huey1999testing; @Havird2020-mj]. Physiological acclimation is driven by endocrine and epigenetic processes that change the underlying physiology to allow organisms to maintain physiological performance around a fitness optimum despite changes in the environment [@seebacher2019histone; @little2013thyroid; @taff2016endocrine]. However, the focus up to now has been primarily on mean physiological responses neglecting how variability in physiological rates might also be impacted at higher temperatures.

>>>>

Phenotypic plasticity is expected to evolve in populations when environmental variability is high but predictable and the costs of plasticity are low [@Chevin2017; @Scheiner2020; @Nunney2016; @Dewitt1998; @Reed2010]. Reversible phenotypic plasticity, such as physiological acclimation, is relatively rapid and can be fine-tuned to proximate environmental conditions making it the first ‘line-of-defense’ against rapid environmental change [@Dewitt1998; @Scheiner2020]. For example, physiological rates are known to speed up as temperature increases because of the thermodynamic effects on chemical reaction rates -- so called 'acute' temperature responses. Acute temperature increases that push organisms from their thermal optimum can be mitigated by plastically adjusting reaction rates in response to chronic temperatures through acclimation [@Seebacher2015-wi; @Havird2020-mj]. Acclimation alters acute thermal sensitivity but does not normally result in physiological rates returning to their original state (‘complete’ compensation - *sensu* @huey1999testing), but rather are only partially compensated [@huey1999testing; @Havird2020-mj]. Physiological acclimation is driven by endocrine and epigenetic processes that change the underlying physiology to allow organisms to maintain physiological performance around a fitness optimum despite changes in the environment [@seebacher2019histone; @little2013thyroid; @taff2016endocrine]. However, the focus up to now has been primarily on mean physiological responses neglecting how variability in physiological rates might also be impacted at higher temperatures.
